---
title: "making fake predators"
author: "Document Author"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---

## making fake predator prey interactions

Here is Ben's model

$$
\begin{align}
F_{ij}^{real} &= \frac{F_{ij}^{max}}{1+h_j*\sum{F_{ij}^{max}}} \\
F_{ij}^{max} &= \alpha_{ij}*N_i*N_j*M_i \\
\alpha_{ij} &= \frac{\phi_{ij}}{D_j}
\end{align}
$$

```{r}

curve(x/(1+.003*x), xlim = c(0,2000))

```

still thinking about the curve

```{r}
curve(x/(1 + 0.4*(15.5)), xlim = c(0,3))
curve(x/(1 + 0.04*(15.5)), xlim = c(0,3), add = TRUE)
curve(x/(1 + 0.8*(15.5)), xlim = c(0,3), add = TRUE)

```

```{r}
freels <- exp(rnorm(10, mean = 2, sd = 1))
freels

# freels = x / (1 + .4*sum(freels))

x = freels * (1 + .4*sum(freels))

plot(x = x, y = freels)


```





NOTE : might correct subscripts

### Flux max

first: one community

```{r}
prey_rich <- 20
pred_rich <- 10

N_prey <- runif(prey_rich, min = .5, max = 1e6)

N_pred <- runif(pred_rich, min = .1, max = 2e5)

# make a list of species names

```


working with pairwise matrices only. 

```{r}
# for each predator we need to know what they eat

prey_names <- paste0("prey", 1:prey_rich)

pred_names <- paste0("pred", 1:pred_rich)

# grab some prey for each predator
prey_names[which(rbinom(prey_rich, 1, .3) == 1)]


library(tidyverse)

fake_community <- tibble(pred_names) |> 
  rowwise() |> 
  mutate(prey_names = list(prey_names[which(rbinom(prey_rich, 1, .3) == 1)])) |> 
  unnest(prey_names) |> 
  mutate(pred_num = parse_number(pred_names),
         prey_num = parse_number(prey_names))

```


Add species level info to this

```{r}

prey_tonne_per_km2 <- runif(N_prey, min = 2e-4, max = 2e5)
pred_tonne_per_km2 <- runif(N_pred, min = 1e-5, max = 9e3)

avg_bodymass_prey <- runif(N_prey, min = 2e-6, .3)
avg_bodymass_pred <- runif(N_pred, min = 2e-6, 14)

# degree isn't actually a random variable =-- more interesting! its missing data
# degree <- rpois(N_pred, lambda = 11.69)

fake_community_data <- fake_community |> 
  # add biomass
  mutate(biomass_pred = pred_tonne_per_km2[pred_num],
         biomass_prey = prey_tonne_per_km2[prey_num],
         bodymass_pred = avg_bodymass_pred[pred_num],
         bodymass_prey = avg_bodymass_prey[prey_num],
         # N_r, r for resouces, so prey
         N_r = biomass_prey/bodymass_prey,
         N_c = biomass_pred/bodymass_pred)



```

simulate these $\phi_{ij}$ numbers

* could be just a random number
* could be correlated for the same prey / same predator

for the second option, something like 

$$
\text{log}(\phi_{ij}) = \alpha_j + \beta_i
$$

but we're not doing that

```{r}

phi_values <- exp(rnorm(nrow(fake_community_data), mean = -2.5, sd = 1))
hist(phi_values)

# every row gets a different phi
fake_comm_phi <- fake_community_data |> 
  mutate(phi_val = phi_values)
```

# simulating flux

_question from Andrew_ does the **MAX FLUX** have any error around it?
or do we assume it is perfectly determined by the equation and only the "F-real" values have error?

We're going to assume it is perfect

the following is **controversial**

```{r}
# calc Fmax
fake_Fmax_sum <- fake_comm_phi |> 
  group_by(pred_names) |> 
  mutate(
    degree = n(),
    Fmax = phi_val/degree * N_r * N_c * bodymass_prey,
    sum_Fmax = sum(Fmax)
  )
```


need handling times

```{r}
handling <- exp(rnorm(N_pred, mean = -2, sd = 2))

fake_Freal <- fake_Fmax_sum |> 
  mutate(h = handling[pred_num],
         Freal = Fmax / (1 + h*sum_Fmax))


```


```{r}
knitr::kable(head(fake_Freal, n = 30))
```


```{r}
fake_Freal |> 
  ggplot(aes(x = Freal)) + geom_histogram()
```

```{r}
fake_Freal |> 
  ggplot(aes(x =  N_r * N_c * bodymass_prey, y = Freal)) + 
  geom_point()
```

```{r}
fake_Freal |> 
  ggplot(aes(x =  N_r * N_c * bodymass_prey, y = Freal)) + 
  geom_point()

```


```{r}
fake_Freal |> 
  group_by(pred_names, sum_Fmax) |> 
  summarise(sum_Freal = sum(Freal))
```



