---
title: "predictions for continuous models"
author: "Andrew"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---

## plotting models with two predictors

```{r, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
library(lme4)
```

plot of the data

```{r}
penguins |> 
  ggplot(aes(y = body_mass_g, x = bill_depth_mm, size = bill_length_mm, colour = species)) + 
  geom_point(alpha = .6) + 
  scale_colour_brewer(type = "qual")
```


predict body mass from the bill sizes

```{r}
mass_model <- lmer(body_mass_g ~ bill_depth_mm + bill_length_mm + (1|species), data = penguins)
summary(mass_model)
```

Center the predictors

```{r}
penguin_centre <- penguins |> 
  mutate(bill_d_c = bill_depth_mm - mean(bill_depth_mm, na.rm = TRUE),
         bill_l_c = bill_length_mm - mean(bill_length_mm, na.rm = TRUE))

```


```{r}
mass_model_c <- lmer(body_mass_g ~ bill_d_c + bill_l_c + (1|species), data = penguin_centre)
```


predictions and put on a figure

```{r}
peng_plot <- penguin_centre |> 
  ggplot(aes(y = body_mass_g, x = bill_d_c, size = bill_l_c, colour = species)) + 
  geom_point(alpha = .6) + 
  scale_colour_brewer(type = "qual")

peng_plot
```

```{r}
library(modelr)

fake_penguins <- data_grid(penguin_centre,
          bill_d_c = seq_range(bill_d_c,n = 30),
          bill_l_c = seq_range(bill_l_c,n = 5),
          species)

fake_penguins$pred <- predict(mass_model_c, newdata = fake_penguins, type = "response")
```

add these to our plot


```{r}
fake_penguins |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + geom_line() + facet_wrap(~bill_l_c) 
```

we get a jagged line between if we omit the other continuous variable

```{r}
fake_penguins |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + geom_point() 
```

We can think of a model with two continuous predictors as representing a smooth 2 dimensional surface:

```{r}
more_fake_penguins <-  data_grid(penguin_centre,
          bill_d_c = seq_range(bill_d_c,n = 30),
          bill_l_c = seq_range(bill_l_c,n = 30),
          species)


more_fake_penguins$pred <- predict(mass_model_c, newdata = more_fake_penguins, type = "response")


more_fake_penguins |> 
  ggplot(aes(x = bill_d_c, y = bill_l_c, fill = pred)) + 
  geom_raster()
```

but this is too hard to visualize! 

## Predicting for one at a time

Here we make two plots: one for bill depth, fixing length at the average. 
Then a second for bill length, fixing depth at the average

This is made easy by centering -- we know the average is now 0

```{r}

fake_penguins_0_length <-  
  data_grid(penguin_centre,
            bill_d_c = seq_range(bill_d_c,n = 30),
            bill_l_c = 0,
            species)


fake_penguins_0_length$pred <- predict(mass_model_c,
                                  newdata = fake_penguins_0_length, type = "response")

fake_penguins_0_length |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + 
  geom_line(size = 2) +
  scale_colour_brewer(type = "qual") + coord_cartesian(ylim = c(2000, 6500)) 
```

```{r}
fake_penguins_0_depth <-  
  data_grid(penguin_centre,
            bill_d_c = 0,
            bill_l_c = seq_range(bill_l_c,n = 30),
            species)


fake_penguins_0_depth$pred <- predict(mass_model_c,
                                  newdata = fake_penguins_0_depth, type = "response")

fake_penguins_0_depth |> 
  ggplot(aes(x = bill_l_c, y = pred, colour = species)) + 
  geom_line(size = 2) +
  scale_colour_brewer(type = "qual")+ coord_cartesian(ylim = c(2000, 6500)) 
```



An alternative is to select only a few values for one predictor. This simplifies the figure and lets us understand the model more completely.

```{r}
# using cut to make a tryptich

range(penguin_centre$bill_l_c, na.rm = TRUE)


fake_penguins_three_lengths <-  
  data_grid(penguin_centre,
            bill_d_c = seq_range(bill_d_c,n = 30),
            bill_l_c = c(-10, 0, 10),
            species)


fake_penguins_three_lengths$pred <- predict(mass_model_c,
                                  newdata = fake_penguins_three_lengths, type = "response")

fake_penguins_three_lengths |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + 
  geom_line(size = 2) +
  facet_wrap(~bill_l_c) + 
  scale_colour_brewer(type = "qual") 
```



### Using group averages to make predictions

The above figure is great, but has one deficiency: it is hard to place the raw data on these figures. 
We can solve this by deriving our three selected values of `bill_l_c` from the data.
To do this, we use `cut()` to divide the column `bill_l_c` into three equal categories. 
In each of these three categories, we calculate the average to use in predictions.
We can then cut the real data in the same three categories. 
This lets us plot the data together with the model. 


```{r}
# using cut to make a tryptich
cut_raw_peng <- penguin_centre |> 
  filter(!is.na(bill_depth_mm)) |> 
  mutate(cut_bill_len = cut(bill_l_c, breaks = 3)) |> 
  group_by(cut_bill_len) |> 
  mutate(mean_bill_len = mean(bill_l_c, na.rm = TRUE)) |> 
  ungroup()


fake_cut_penguins <-  
  data_grid(cut_raw_peng,
            bill_d_c = seq_range(bill_d_c,n = 30),
            nesting(bill_l_c = mean_bill_len,
                    cut_bill_len),
            species)


fake_cut_penguins$pred <- predict(mass_model_c, newdata = fake_cut_penguins, type = "response")

fake_cut_penguins |> 
  count(bill_l_c, cut_bill_len)


fake_cut_penguins |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + 
  geom_line(size = 2) +
  facet_wrap(~cut_bill_len) + 
  scale_colour_brewer(type = "qual") + 
  geom_point(aes(x = bill_d_c, y = body_mass_g), data = cut_raw_peng, alpha = .6)
  
```

### interactions and better models

```{r}
mass_model_interact <- lmer(body_mass_g ~ bill_d_c*bill_l_c + (1 + bill_d_c*bill_l_c|species), data = penguin_centre)
summary(mass_model_interact)

# using cut to make a tryptich
cut_raw_peng <- penguin_centre |> 
  filter(!is.na(bill_depth_mm)) |> 
  mutate(cut_bill_len = cut(bill_l_c, breaks = 3)) |> 
  group_by(cut_bill_len) |> 
  mutate(mean_bill_len = median(bill_l_c, na.rm = TRUE)) |> 
  ungroup()


fake_cut_penguins <-  
  data_grid(cut_raw_peng,
            bill_d_c = seq_range(bill_d_c,n = 30),
            nesting(bill_l_c = mean_bill_len,
                    cut_bill_len),
            species)


fake_cut_penguins$pred <- predict(mass_model_interact, newdata = fake_cut_penguins, type = "response")

fake_cut_penguins |> 
  count(bill_l_c, cut_bill_len)


fake_cut_penguins |> 
  ggplot(aes(x = bill_d_c, y = pred, colour = species)) + 
  geom_line(size = 2) +
  facet_wrap(~cut_bill_len) + 
  scale_colour_brewer(type = "qual") + 
  geom_point(aes(x = bill_d_c, y = body_mass_g), data = cut_raw_peng, alpha = .6)
```

group mean centered

```{r}
penguins_gc <- penguins |> 
  group_by(species) |> 
  mutate(bill_len_gm = mean(bill_length_mm, na.rm = TRUE),
         bill_len_gc = bill_length_mm - bill_len_gm) |> 
  ungroup()

bill_model <- lmer(body_mass_g ~ bill_len_gm + bill_len_gc + (1 + bill_len_gc | species), data = penguins_gc)

fake_gc_peng <- penguins_gc |> 
  nest_by(species, bill_len_gm) |> 
  mutate(bill_len_gc = list(seq_range(data$bill_len_gc, n = 12))) |> 
  select(-data) |> 
  unnest(bill_len_gc) |> 
  ungroup()

fake_gc_peng$pred <- predict(bill_model, newdata = fake_gc_peng, type = "response")

fake_gc_peng |> 
  ggplot(aes(x = bill_len_gm + bill_len_gc, y = pred, colour = species)) + 
  geom_line(size = 2) + 
  scale_colour_brewer(type = "qual") + 
  geom_point(aes(y = body_mass_g), data = penguins_gc)
```

