---
title: "constant or linear decline in the rate"
author: "Vignette Author"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
curve(exp(-plogis(-3 + .6*x)*x), xlim = c(0, 10))
curve(exp(-plogis(3 - 1.2*x)*x), xlim = c(0, 10))
curve(exp(-x), xlim = c(0, 10))
curve(exp(-.5*x), xlim = c(0, 10))

```

yeah its 1 - exp(-exp(x))

```{r}
curve(1 - exp(-exp(x)), xlim = c(-3, 3))
curve(exp(-exp(x)), xlim = c(-3, 3))
curve(1 - exp(-exp(-x)), xlim = c(-3, 3))
curve(1 - exp(-exp(-1)*x), xlim = c(-3, 3))
```

What is the for the probabiligy takes as it goes down? 

if you can do this with the cloglog link, then you can spare yourself effort and use offset (once again) to look at different functions of decline

```{r}
curve(exp(-exp(x)), xlim = c(-3, 3))
curve(exp(-exp(x)*x), xlim = c(-3, 3))
#weird curve!



```
what a weird curve! idk if that's what I wanted

```{r}
curve(exp(-.1 + .5 * x), xlim = c(0, 3))


```

pielou equation is 

```{r}
curve(exp(-.7*x), xlim = c(0,4))
```
and there x is time.

```{r}
curve(exp(-1.5*x), xlim = c(0,4))
```

So it can really be any negative number -- doesn't have to be between 0 and 1

```{r}
curve(5*x), xlim = c(0,4))
```

```{r}
curve(exp(-exp(-1.3)*x), xlim = c(0,3), ylim = c(0, 1))
```
So the really wild thing to note is that this is just reprasing the constant exponential decline model of pielou. it looks just like the cloglog function. Time acts as the exponent (rules of exponents)
But I feel like in R you can model this 

```{r}
lnk <- binomial(link = "logit")
curve(lnk$linkfun(x), xlim = c(0, 1))
curve(lnk$linkinv(x), xlim = c(-3, 3))

curve(binomial(link = "cloglog")$linkinv(x), xlim = c(-3, 3))


```

```{r}

curve(exp(-exp(-1.3)*x), xlim = c(0,3), ylim = c(0, 1))
```


```{r}

curve(binomial(link = "cloglog")$linkinv(-2 + log(x)), xlim = c(0, 30))

```

Doi the binomial, but reverse the number of failures and the number of successes at each turn

can you do with with the Normal link?

```{r}
gaussian(link = "cloglog")
```

For the proportion gone of a gaussian quantity. 

well, if we know the precise quantities, then yes. you just calculate the proportion and use, e.g., the normal distribution. you might choose to get fancy and us the beta distribution

model tadpole growth in this way? considering the growth rate and the max size and correlated reponses? is there wa way to put max size into this expression? I kinda don't think so

Is ther ea way to model logistic population growth with the logistic function? really seems like there should be


 
```{r}
curve(binomial(link = "cloglog")$linkinv(-7 + 1*x + log(x)), xlim = c(0, 30))
```
 
 but it is the complement of that
 
```{r}
curve(1 - binomial(link = "cloglog")$linkinv(-7 + 1*x + log(x)), xlim = c(0, 30))
```

or something?? 

just messing around.. an you scale time??
```{r}
curve(binomial(link = "cloglog")$linkinv(-7 + 1*x + log(x) - .4*log(x)), xlim = c(0, 30))
```


# could make it truly insane and use the log link for the slope on time. it basicaly has to decline, which means that the probaility gets closer and closer to 1. ie the 


```{r}
curve(exp(-x), xlim = c(0,1))
```

```{r}
curve(exp(x), xlim = c(-1,1))
abline(v = 0)

```

so it can actually be any negative real number -- doesn't have to be between 1 and 0. Or does it, for population reasons? this way it bascially has an asymptote at 1/e, or `exp(-1)`

```{r}
curve(exp(-plogis(-x)*x), xlim = c(0, 10))

xs <- seq(0,10,length.out = 45)
ys <- exp(-plogis(-xs))
plot(ys[-length(ys)] - ys[-1])
```

This function seems to have a little kick at the beginning that I think looks like it really works. Like it looks very interesting, and it could mabye be a way to model soil decompositon also.


ANyway I think it works out to being basically the Gompertz curve
