---
title: "Simulating tadpoles"
author: "Andrew and Simon"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---

## genetic information

cluster -- is one egg mass with multiple fertilizing males

family -- is tadpoles fertilizes by the same mother-father pair

genetics applies only to the final size of specific indivituals


* we're removing the LOW herbicide treatment because 

* trying to catch before legs come out BUT sometimes

svl is not always correlated to mass -- because over the time int he final stage of life, mass decreases per time 


## simulation of Tadpole final size

start with the mass of 20 tadpoles
```{r}
# twenty tadpoles
tadpole_mass <- rnorm(n = 20, mean = .8, sd = .2)

```

might not actualy be normal

```{r}
hist(tadpole_mass)
```


now add the effect of drying

```{r}
avg_healthy_good_water <- .8
effect_dry <- c("null" = 0, "medium" = -.03, "rapid" = -0.1)

avg_healthy_good_water + effect_dry["medium"]

library(tidyverse)

fake_data <- tibble(dry_treatment = rep(names(effect_dry), each = 20),
                    avg_size = avg_healthy_good_water + effect_dry[dry_treatment],
                    obs_size = rnorm(length(dry_treatment), mean = avg_size, sd = .02)) |> 
  mutate(dry_treatment = factor(dry_treatment, levels = c("null", "medium", "rapid")))

ggplot(fake_data, aes(x = dry_treatment, y = avg_size)) + geom_point()

fake_data |> 
  ggplot(aes(x = dry_treatment, y = obs_size)) + 
  geom_boxplot()+
  geom_point()


```

Now we try to recover those values! 

```{r}
summary(lm(obs_size ~ dry_treatment, data = fake_data))
```

### adding poison

```{r}

```


### adding random effects

```{r}
# how many cattle tanks?

n_cattle_tanks <- 12
sd_cattle_tanks <- .03

n_bins <- 24 #(two bins in a tank)
sd_bins <- .1

avg_healthy_good_water <- .8
effect_dry <- c("null" = 0, "medium" = -.03, "rapid" = -0.1)

effect_poison <- c("null" = 0, "faible" = -.01, "moyen" = -.1,"haut" = -.3)

tank_effects <- rnorm(n_cattle_tanks, mean = 0, sd = sd_cattle_tanks)

bin_effects <- rnorm(n_bins, mean = 0, sd = sd_bins)

# sd(bin_effects); sd_bins
# sd(tank_effects); sd_cattle_tanks

treatment_combos <- expand.grid(
  dry_trt = names(effect_dry),
  poison_trt = names(effect_poison)
)

# repeat every row twice! 
treatments_for_bins <- treatment_combos[rep(1:12, each = 2),]

fake_data_avgs <- tibble(
 tank_id = rep(1:n_cattle_tanks, each = 2) ,
 bin_id  = 1:n_bins,
 tank_effect = tank_effects[tank_id],
 bin_effect = bin_effects[bin_id],
 not_controlled_effect = tank_effect + bin_effect
) |> 
  cbind(treatments_for_bins) |> 
  # add the effects of treats 
  mutate(avg_effect =  avg_healthy_good_water + effect_dry[dry_trt] + effect_poison[poison_trt] + not_controlled_effect)

knitr::kable(fake_data_avgs)
```


make observations of the tadpoles

(begin: tidyverse magick! )

```{r}
n_metamorphs <- 20
sd_mass <- .1

fake_full_data <- fake_data_avgs |> 
  rowwise() |> 
  mutate(obs_mass = list(rnorm(n = n_metamorphs, mean = avg_effect, sd = sd_mass))) |> 
  unnest(obs_mass)
```

Try to model it! 

```{r}
library(lme4)

fm <- lmer(obs_mass ~ dry_trt + poison_trt + (1 | tank_id) + (1 | bin_id), data = fake_full_data)
summary(fm)
```

```{r}
fixef(fm)
effect_dry
effect_poison
avg_healthy_good_water
```

```{r}
ranef(fm)
bin_effects
```

```{r}
summary(fm)$varcor

sd_bins
sd_cattle_tanks
```


## plot the fake data

```{r}
fake_full_data |> 
  ggplot(aes(x = poison_trt, y = obs_mass)) + 
  geom_violin() + 
  facet_wrap(~dry_trt)
```

```{r}

fake_full_data |> 
  ggplot(aes(x = poison_trt, y = obs_mass)) + 
  geom_violin() + 
  facet_grid(tank_id~dry_trt)
```

Why the singularity in Simon's model?

```{r}
fake_full_data |> 
  count(tank_id, dry_trt)
```

This is just as it should be! one dry trt is applied to each tank
